\textbf{ Accurate spectrophotometric estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-31 paper/sp.tex
apogee data and cuts

One final detail: We perform all of the fitting in a two-fold train-and-test framework,
in which the data are split into two disjoint subsets, A and B.
The model trained on the A data is used to predict the $Q_m$ values (and hence
the luminosities, distances, and parallaxes) of the B data,
and the model trained on the B data is used to 
predict the $Q_m$ values of the A data.
This ensures that, in the estimate of any individal star's luminosity, distance,
or parallax, none of the \gaia\ data pertaining to that star was used.
This makes the parallax estimates from the spectrophotometric feature vectors
$x_m$ statistically independent of the \gaia\ parallax measurements.
They are not independent \emph{globally}, but they are independent on a
star-by-star basis.

@@

ACE: We take the \apogee\ data from what subset?
We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.

>>>>

One final detail: We perform all of the fitting in a two-fold train-and-test framework,
in which the data are split into two disjoint subsets, A and B.
The model trained on the A data is used to predict the $Q_m$ values (and hence
the luminosities, distances, and parallaxes) of the B data,
and the model trained on the B data is used to 
predict the $Q_m$ values of the A data.
This ensures that, in the estimate of any individal star's luminosity, distance,
or parallax, none of the \gaia\ data pertaining to that star was used.
This makes the parallax estimates from the spectrophotometric feature vectors
$x_m$ statistically independent of the \gaia\ parallax measurements.
They are not independent globally---\gaia\ data was used to train the model---but
each star's spectrophotometric parallax estimate is independent
of the astrometric estimate from \gaia\ on a star-by-star basis.

@@

We take the \apogee\ (HOGG CITE) spectral data
from \sdssiv\ (HOGG CITE) \acronym{DR14} (HOGG CITE).
Because we want to make a purely linear model, we restrict our consideration
to a small region in stellar parameter space.
We cut the \apogee\ data down to the range $0<\logg<2.2$, which isolates
stars that are more luminous than the red clump.
The \apogee\ pipeline $\logg$ values (which we use) have uncertainties but
this cut leads to a clean sample of luminous red giants, and it is a cut
that is only on spectral properties of the stars (and not photometry nor astrometry).
